---
title: "ISLR Lab 4: Classification"
output: html_notebook
---
```{r echo = FALSE, message = FALSE}
library(ISLR)
library(tidyr)
library(purrr)
library(dplyr)
library(ggplot2)
library(caret)
library(GGally)
```

We will begin by examining some numerical and graphical summaries of the `Smarket` data, which is part of the 
`ISLR` library. 

```{r}
glimpse(Smarket)
head(Smarket)
cor(Smarket[, -9]) %>%
  ggcorr(label = TRUE)
```
Next, we will fit a logistic regression model in order to predict `Direction` using `Lag1` through `Lag5` and
`Volume`.
```{r}

caret_lr <- train(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume,
                  data = Smarket,
                  method = 'glm',
                  family = 'binomial')
summary(caret_lr)
```
The `predict()` function can be used to predict the probability that the market will go up, given values of the 
predictors. The type="prob" option tells `caret` to output probabilities of the form $P(Y = 1|X)$, as opposed to 
other information such as the logit. If no data set is supplied to the `predict()` function, then the probabilities
are computed for the training data that was used to fit the logistic regression model. 
```{r}
predicted_probabilities <- predict(caret_lr, type = 'raw')
head(predicted_probabilities)
```
```{r}
table(predicted_probabilities, Smarket$Direction)
```
```{r}
train <- Smarket %>%
  filter(Year <2005)
test <- Smarket %>%
  filter(Year == 2005)

model_data_split <- train(Direction ~ Lag1 + Lag2 + Lag3 + Lag4 + Lag5 + Volume,
                          method = 'glm',
                          family = 'binomial',
                          data = train)
new_predictions <- predict(model_data_split, test)
mean(new_predictions != test$Direction)

```
LDA:
```{r}
lda_model <- train(Direction ~ Lag1 + Lag2, 
                   method = 'lda',
                   data = train)
confusionMatrix(test$Direction, predict(lda_model, test))
lda_model$finalModel
predictions <- predict(lda_model, test)
predictions_prob <- predict(lda_model, test, type = 'prob')
table(predictions, test$Direction)
mean(predictions == test$Direction)
sum(predictions_prob[,1] > .9)
```
QDA
```{r}
qda_model <- train(Direction ~ Lag1 + Lag2, 
                   method = 'qda',
                   data = train)
predict_qda <- predict(qda_model, test)
mean(predict_qda == test$Direction)
```
KNN
```{r}
knn_model <- train(Direction ~ Lag1 + Lag2,
                   method = 'knn',
                   data = train,
                   preProcess = c('center', 'scale'))
predict_knn <- predict(knn_model, test)
table(predict_knn, test$Direction)
mean(predict_qda == test$Direction)
```
APLICACIÓN: usando caret
```{r}
nzv <- nearZeroVar(Caravan)
caravan_clean <- Caravan[,-nzv]
split_index <- createDataPartition(caravan_clean$Purchase, p = 0.75, list = FALSE)
train_crv <- caravan_clean[split_index,]
test_crv <- caravan_clean[-split_index,]
ctrl <- trainControl(method="repeatedcv",repeats = 3)
knn_crv <- train(Purchase ~ .,
                 data = train_crv,
                 method = 'knn',
                 preProcess = c('center', 'scale'),
                 trControl = ctrl,
                 tuneLength = 15)
predictions_crv <- predict(knn_crv, test_crv)
table(predictions_crv, test_crv$Purchase)
mean(predictions_crv == test_crv$Purchase)

```

APLICACIÓN: SIN USAR CARET
```{r}
##Ahora con algo clásico
library(class)
train_crv_nc <-  train_crv %>%
  select( -Purchase) %>%
  scale()
test_crv_nc <- test_crv %>%
  select(-Purchase) %>%
  scale()

knn_crv_nc_1 <- knn(train = train_crv_nc, test = test_crv_nc, cl = train_crv$Purchase, k = 1)
(t1 <- table(knn_crv_nc_1, test_crv$Purchase))
mean(knn_crv_nc_1 == test_crv$Purchase)
t1[4] / (t1[4]+ t1[2])

knn_crv_nc_3 <- knn(train = train_crv_nc, test = test_crv_nc, cl = train_crv$Purchase, k = 3)
(t3 <- table(knn_crv_nc_3, test_crv$Purchase))
mean(knn_crv_nc_3 == test_crv$Purchase)
t1[4] / (t1[4]+ t1[2])

knn_crv_nc_5 <- knn(train = train_crv_nc, test = test_crv_nc, cl = train_crv$Purchase, k = 5)
(t5 <- table(knn_crv_nc_5, test_crv$Purchase))
mean(knn_crv_nc_5 == test_crv$Purchase)
t5[4] / (t5[4]+ t5[2])
```

