---
title: "ISLR6 - Lab 1: Feature selection "
output: html_notebook
author: Jorge Rotter
---

This lab on Subset Selection in R comes from p. 244-247 of "Introduction to Statistical Learning with 
Applications in R" by Gareth James, Daniela Witten, Trevor Hastie and Robert Tibshirani. It was, however, 
rewritten for using the `caret` package. It was adapted for use of the `caret` package.

```{r echo = FALSE, warning = FALSE}
library(ISLR)
library(plyr)
library(caret)
library(dplyr)

```


#Introduction

Here we apply the best subset selection approach to the Hitters data. We wish to predict a baseball player’s 
Salary on the basis of various statistics associated with performance in the previous year. Let's take a quick 
look:

```{r warning=FALSE, echo = FALSE}
library(ISLR)
library(caret)
library(dplyr)
head(ISLR::Hitters)
```

First of all, we note that the `Salary` variable is missing for some of the players. The `is.na()` function can 
be used to identify the missing observations. It returns a vector of the same length as the input vector, with a 
`TRUE` value for any elements that are missing, and a `FALSE` value for non-missing elements.
The `sum()` function can then be used to count all of the missing elements:

```{r}
sum(is.na(ISLR::Hitters$Salary))
```

We see that `Salary` is missing for 59 players. The `na.omit()` function removes all of the rows that have
missing values in any variable. Alternatively, caret can do median imputation or knn imputation (which of course
we are not doing, since `Salary` is our response.)

```{r}
# Print the dimensions of the original Hitters data (322 rows x 20 columns)
dim(ISLR::Hitters)

# Drop any rows the contain missing values
Hitters <- na.omit(ISLR::Hitters)

# Print the dimensions of the modified Hitters data (263 rows x 20 columns)
dim(Hitters)

# One last check: should return 0
sum(is.na(Hitters))
```

#Recursive feature elimination (Backwards Selection)
`caret` has several algorithms for feature selection. We'll go through RFE first. Caret has two functions for
performing RFE. `rfeIter` and `rfe`. Although strictly speaking `rfeIter` is the algorithm covered by the book,
the `rfe` function includes a cross-validation approach to simplify what is done in pages 250-251. As such, we'll
focus in `rfe` only.

We first create a control object to specify that we wish to use 10-fold cross validation (5 times) to perform 
variable selection in a linear model (hence the `lmFuncs`. There are also prebuilt `rfFuncs`, `nbFuncs`, 
`treebagFuncs` and `caretFuncs`).

```{r}
set.seed(1)
rfe_control <- rfeControl(functions = lmFuncs,
                          method = "repeatedcv",
                          number = 10,
                          repeats = 5,
                          verbose = FALSE)

training_index <- createDataPartition(Hitters$Salary, p = 0.75, list = FALSE)
prim_index_train <- sample(c(TRUE,FALSE), nrow(Hitters),rep=TRUE)
prim_index_test <- !prim_index_train

training_set <- Hitters[traino,] %>%
  select(-Salary)
training_set_mm <- model.matrix(~., data = training_set)[,-1]

test_set <- Hitters[!traino,] %>% 
  select(-Salary)
test_set_mm <- stats::model.matrix(~., data = test_set)[,-1]

training_values <- Hitters[traino,]$Salary

model <- rfe(x = training_set_mm, 
    y = training_values,
    sizes = c(1:ncol(training_set_mm)),
    rfeControl = rfe_control)
model
```

#Genetic algorithms
```{r}
gafs_control <- gafsControl(functions = caretGA ,
                    method = 'repeatedcv',
                    number = 2,
                    repeats = 2,
                    verbose = TRUE)
model_ga <- gafs(x = training_set_mm,
                 y = training_values,
                 iters = 2,
                 gafsControl = gafs_control,
                 method = 'lm')
```

